\propertitle

### 2019-11-21 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-20T20:23:20+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The number and degrees of the constraining moments depend on the questions
and hypotheses that a researcher is exploring; for example, hypotheses
about the \enquote{cooperativity} or \enquote{interaction} in the network
activity. We discuss this kind of use in \sect~\ref{sec:marginalization}.
Note that the constraint equation~\eqref{eq:final_constraint} for the
distribution~\eqref{eq:app_maxent_pop_first} \emph{may not have exact
  solutions} if $N$ is strictly larger than $n$; and the possible
discrepancy increases with the number of constraints. The essential reason
for this discrepancy is the approximate character of two assumptions on
which the maximum-entropy method is based and of one assumption specific to
our application: measured moments equal expectations (or equivalently that
the number of time bins $T$ is infinite), the order of the time bins is
irrelevant, and that all full-network subsets with $\yA$ neurons are
equally likely to be generating a total activity $\yA$. We discuss these
assumptions in \sect~\ref{sec:assumptions}.

@@

Before applying the formula above to two concrete data sets we want to add
three remarks on the maximum-entropy method that are seldom made in the
neuroscientific literature. They are important for the interpretation of
the results and are further discussed in \sect~\ref{sec:assumptions}.

First, a maximum-entropy distribution, like $\px(\yA \| M, N)$, is the
zeroth-order approximation (in the sense of Laplace's
approximation)\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}
of four different distributions for the full network, which differ
numerically from one another in higher-order approximations:
\begin{enumerate}[label=(\roman*)]
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{enumerate}

Second, the maximum-entropy method % (based on the \emph{Shannon} relative
% entropy, as opposed to other entropy functions such as Burg's
% \citeyear{burg1975})
rests on some implicit assumptions about the probabilities for the long-run
frequency distribution of activities. The often heard statement that it's
\enquote{the maximally unbiased (or non-committal) distribution} must be
taken with a grain of salt. \citep[\mynote{move to
  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}.

The number of moments used with this method depends on the questions and
hypotheses that a researcher is exploring; for example, hypotheses about
the \enquote{cooperativity} or \enquote{interaction} -- for example,
pairwise or higher-order -- of the network activity. We discuss this kind
of use in
\sect~\ref{sec:marginalization}.%, but our main goal is to simply introduce
% the general method without entering into biological questions of this
% kind.

The particular case in which $n$ moments are constrained is especially
important: it corresponds to fully constraining the marginal frequency
distribution for the activity of the sample neurons, $\yf$. In this case,
our belief about the full-network activity is based on all available
measured frequency data. Note that the application of the maximum-entropy
method \emph{at the sample level} is trivial and meaningless in this case
-- it just gives back the measured frequency distribution. But application
of the method \emph{at the level of the full network} is not trivial.

Using the full frequency distribution of the sample may be a bad idea,
however, because the maximum-entropy distribution may become a bad
approximation of some of the four distributions described above. It is
preferable to use a moderately high number of moments smaller than $n$. We
explain this point in \sect***.

@@

Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
log-ratio of the probabilities of the data $\yf$ conditional on the
hypotheses $M''$ and $M'$:
\begin{equation}
  \label{eq:bayes_factor}
  \ydi(M'',M') =
  \log\bigl[\pf(\yf \| M'')/
  \pf(\yf \| M')\bigr].
\end{equation}
This is called their \emph{relative weight of evidence}, the logarithm of
their \emph{relative Bayes factor}
\citep[\chap~6]{good1950}{good1975,good1981,good1985,good1983}[\sect~1.4]{osteyeeetal1974}{mackay1992,kassetal1995}[see
also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
this equality in appendix**. So the exponential of $\ydi(M'',M')$ tells us
how much more probable the data $\yf$ are conditional on $M''$, than
conditional on $M'$. We can also combine this measure with pre-data
probabilities for the two hypotheses to obtain the ratio of their
probabilities conditional on the data.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-21T11:47:50+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The number and degrees of the constraining moments depend on the questions
and hypotheses that a researcher is exploring; for example, hypotheses
about the \enquote{cooperativity} or \enquote{interaction} in the network
activity. We discuss this kind of use in \sect~\ref{sec:marginalization}.
Note that the constraint equation~\eqref{eq:final_constraint} for the
distribution~\eqref{eq:app_maxent_pop_first} may not have exact solutions
if $N$ is strictly larger than $n$; and the possible discrepancy typically
increases with the number of constraints. The essential reason for this
discrepancy is the approximate character of two assumptions behind the
maximum-entropy and of one assumption specific to our application: measured
moments equal expectations (or equivalently, the number of time bins $T$ is
infinite), the order of the time bins is irrelevant, and an activity level
$\yA$ can be equally likely be generated by any set of $\yA$ neurons in the
full network. The magnitude of this discrepancy can be a signature of the
(at least temporary) presence of a \enquote{neuronal assembly}
\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
in \sect~\ref{sec:assumptions}.

@@

Before applying the formula above to two concrete data sets we want to add
two remarks on the maximum-entropy method that are seldom made in the
neuroscientific literature. They are important for the interpretation of
the results and are further discussed in \sect~\ref{sec:assumptions}.
First, a maximum-entropy distribution, like $\px(\yA \| M, N)$, is the
zeroth-order approximation (in the sense of Laplace's
approximation)\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}
of four different distributions for the full network, which differ
numerically from one another in higher-order approximations:
\begin{enumerate}[label=(\roman*)]
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{enumerate}
The validity of the maximum-entropy approximation decreases as the ratio
$N/T$ increases. Second, the maximum-entropy
method % (based on the \emph{Shannon} relative
% entropy, as opposed to other entropy functions such as Burg's
% \citeyear{burg1975})
rests on some implicit assumptions about the probabilities for the long-run
frequency distribution of activities, besides the assumptions mentioned at
the end of the previous paragraph. So the often heard statement that it
gives \enquote{the maximally unbiased (or non-committal) distribution} must
be taken with a grain of salt. \citep[\mynote{move to
  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}.

@@

Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
log-ratio of the probabilities of the data $\yf$ conditional on the
hypotheses $M''$ and $M'$:
\begin{equation}
  \label{eq:bayes_factor}
  \ydi(M'',M') =
  \log\bigl[\pf(\yf \| M'')/
  \pf(\yf \| M')\bigr].
\end{equation}
This is called their \emph{relative weight of evidence}, the logarithm of
their \emph{relative Bayes factor}
\citep[\chap~6]{good1950}{good1975,good1981,good1985,good1983}[\sect~1.4]{osteyeeetal1974}{mackay1992,kassetal1995}[see
also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
this equality in appendix~\ref{sec:measure_suff}. So the exponential of
$\ydi(M'',M')$ tells us how much more probable the data $\yf$ are
conditional on $M''$, than conditional on $M'$. We can also combine this
measure with pre-data probabilities for the two hypotheses to obtain the
ratio of their probabilities conditional on the data
\citep[\cf][]{bretthorst2013}.
